Forecasting the Support of Kamala Harris in 2024 U.S. Presidential Election using a Model-Based Approach

### 2024-11-04 paper/ForecastingtheSupportofKamalaHarrisin2024U.S.PresidentialElection.qmd
update paper with abstract

---
title: "Forecasting the Support of Kamala Harris in 2024 U.S. Presidential Election using a Model-Based Approach with Poll-of-Polls Data"
subtitle: "Kamala Harris expects to win on 47.5%, leads Donald Trump by about 1% in 2024 US election"
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This study models the 2024 U.S. Presidential Election by analyzing polling data with adjustments for factors like recency and sample size, simulating an approach that captures up-to-date voter sentiment. Using the recent 4 months (July 21th - November 2th) of aggregated poll data, the model identifies shifts in candidate support across key battleground states, providing a nuanced view of evolving preferences as election day approaches. By emphasizing recent, high-quality polls, the analysis improves predictive reliability, offering an understanding that could inform campaign strategies and public discourse. These findings highlight the value of weighting and time-sensitive adjustments in forecasting methods, balancing accuracy with responsiveness to the latest trends."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, Michigan, and Wisconsin by Model Prediction (From August 1, 2024, to November 2, 2024)."
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)
late<-combined_predictions_filtered[which.max(combined_predictions_filtered$end_date), ]
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate))+
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 0.5)+
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1, size =6),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

## weighetd
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 0.5)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight\nModel: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1,size=6),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```

>>>>

---
title: "Forecasting the Support of Kamala Harris in 2024 U.S. Presidential Election using a Model-Based Approach"
subtitle: "Kamala Harris expects to win on 47.5%, leads Donald Trump by about 1% in 2024 US election in Popular Vote based on Poll-of-Polls Data and Linear Modeling."
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This study models the 2024 U.S. Presidential Election by analyzing polling data with adjustments for factors like recency and sample size, simulating an approach that captures up-to-date voter sentiment. Using the recent 4 months (July 21th - November 2th) of aggregated poll data, we built a multiple linear regression model to estimate Harris’ support in popular vote and the winner of seven key battleground states. We estimate that  Kamala Harris will lead with an expected 47.5% support, ahead of Donald Trump by approximately 1% in popular vote and winning in three of seven key battleground states, Wisconsin, Michigan, and Pennsylvania.  By emphasizing recent, high-quality polls, the analysis improves predictive reliability, offering an understanding that could inform campaign strategies and public discourse. These findings highlight the value of weighting and time-sensitive adjustments in forecasting methods, balancing accuracy with responsiveness to the latest trends."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, Michigan, and Wisconsin by Model Prediction (From August 1, 2024, to November 2, 2024)."
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)
late<-combined_predictions_filtered[which.max(combined_predictions_filtered$end_date), ]
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate))+
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted),se = FALSE, size = 0.5)+
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1, size =6),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

## weighetd
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted),se = FALSE, size = 0.5)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight\nModel: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1,size=6),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```
